\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-04 paper/paper.tex
Updates.

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

@@

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. This is due to the fixed cost of a round trip to the SSD through the IO
interface (NVMe in our case), regardless of the volume of data being
synchronously persisted. Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are significantly more expensive than standard DIMMs. This means that they will
likely be available in small quantities. We compensate for this by periodically
draining the NVDIMM log into a log placed on a standard SSD on the system.

The second key contribution of this work is in showing that the throughput
overheads in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum - a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory eliminating block IO from the persistence step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

The rest of the paper is organized as follows. We describe Cyclone's system
architecture in Section~\ref{sec:sysarch}. We describe this two level log
structure in Section~\ref{sec:storage}. The set of optimizations to the network
stack is decribed in Section~\ref{sec:network}. Both the storage (NVM, SSD) as
well as the NIC expose significant parallelism that is left unused when Cyclone
is used to replicate a single log. To exploit this parallelism we use Cyclone to
replicate multiple logs, each using a different instance of the RAFT consensus
protocol. We show how to exploit the commutativity inherent in key-value store
interfaces to make use of this parallelism in Cyclone's design in
Section~\ref{sec:parallelism}.  A detailed evaluation of Cyclone's replication
performance is provided in the context of RocksDB - a popular persistent key
value store - in Section~\ref{sec:evaluation}. We then discuss related work
before concluding.

@@

\begin{itemize}
\item GET(K): returns value corresponding to key K
\item PUT(K, V): sets the value of key K to value V
\end{itemize}

A key value store typically serves multiple client requests concurrently and is
linearizable~\cite{linearizability}. This means any history of requests and
responses can be rearranged such that responses immediately follow corresponding
requests and if a response precedes a request in the original history it also
does so in the rearranged history. Key value stores such as RocksDB also
support atomic writes to multiple keys and snapshots.

@@

Cyclone uses RAFT~\cite{raft} to keep the logs in sync across the replicas. On a
failure of the leader replica a new leader is automatically elected - RAFT
ensures that the new leader has the most up to date log. The Cyclone client
library automatically locates the new leader on a failover - clients therefore
see only a brief interruption in service. Although read requests (GET
operations) are not logged, they are serviced in FIFO order with the write requests
(PUT operations) currently in Cyclone's log at the moment the read request is
received. This ensures linearizability in the face of failovers.

@@

The two level log design in Cyclone lays the foundation for improvements to the
network stack that we discuss next - a significant contributor to the overheads
in Figure~\ref{fig:problem}. We use RAFT to replicate the NVM log from the
leader replica to the followers. Each replica \emph{independently} drains
committed (in terms of RAFT consensus) entries from the NVM log to the
flashlog. Block IO therefore is no longer a consideration when optimizing data
movement over the network, unlike systems that need to deal with such
problems~\cite{reflex}. Replication reduces to efficiently sending a block of
data already present in directly attached memory over the network.

@@

The core network operation in replication is to receive a request from the
client at the leader replica and send that exact same request (as a log entry)
to all follower replicas. The leader therefore \emph{multicasts} a received
request packet to follower replicas, a problem well studied in the networking
community when building software packet switches. Software packet switches reach
impressive speeds of millions of packets forwarded per second~\cite{dpdk_perf}, a
number far in excess of the few hundreds of packets we manage in
Figure~\ref{fig:problem}. We therefore proceed by approximating a software
packet switch for the networking component of Cyclone. However software packet
switching does not need to deal with failure - usually requiring only a simple
decision for each packet at the node. RAFT introduces significant
protocol level work for each forwarded packet a challenge that we need to deal
with to approach the desired software packet switching efficiency in the common
case.

@@

We direct DPDK to use pages backed by NVM for packet buffers. DPDK uses a
concurrent memory allocator based on reference counting - used by both the NIC
driver and CPU cores. This means that packet allocation and deallocation does
not happen in the same order as their corresponding position in the replicated
RAFT log. To deal with this, we use a level of indirection as shown in
Figure~\ref{fig:nvm_log}. The NVM log is maintained as a circular log of fixed
sized pointers. Adding a level of indirection in the NVM log allows us to
separate the FIFO ordered circular log being manipulated by RAFT from packet
data being managed by the memory allocator of DPDK. Both the circular log and
packet data are in NVM. An advantage of this scheme is that it makes recovery
from NVM easy - appends to the circular pointer log are atomic and we can use
the pointer log to recover allocator state i.e. what pieces of NVM are currently
in use by the log.

@@

\subsection{Eliminating Data Movement}
\label{sec:dm}
Software packet switches built on DPDK try to touch as little of the packet data
as possible, minimizing movement of data up and down the cache and memory
hierarchy. Log replication looks like multicast that does not require deep
packet inspection. Software packet switches implementing multicast therefore
simply manipulate packet headers to produce new packets to send on the output
ports. We designed our implementation of raft around the same principle.

@@

We now turn our attention to the persistence step in
Figure~\ref{fig:control_plane}. RAFT requires that the log entry be persisted
before it is multicast out to follower replicas. DPDK userspace NIC drivers
operate in DDIO mode~\cite{ddio} where the packet is directly written into the
CPU cache rather than first being DMAed into DRAM and then fetched by the CPU on
demand. Since the NVM is directly attached we need to persist the packet to it
by executing a cacheline flush ({\tt clflush}) instruction for every cacheline
in the packet and the pointer in the pointer buffer to persist these via the
memory bus. This is not too onerous a burden because we can use the newly
introduced {\tt clflush-opt}~\cite{clflush_opt} instruction specifically
intended to efficiently flush to persistent memory without the overhead of the
serialization normally introduced by {\tt clflush}. This allows us to hit full
memory bandwidth on present generation platforms, a quantity in excess of 200
Gb/s per core, well above the near term speeds of network interface cards. We
execute a single serializing {\tt sfence} before the sequence of {\tt clflushes}
to make sure any dirty cacheline due to header manipulation related writes from
the CPU are sent to cache.

@@

Finally, we discuss our choice of RAFT as a consensus protocol. A key advantage
of RAFT is that unlike alternatives such as MultiPaxos~\cite{multipaxos} or
Quorum based replication~\cite{quorum}, the leader is guaranteed to have the
most up to date logs. There is therefore no case where a leader needs to receive
log entries to ``fill'' holes in its logs from follower replicas, simplifying
the protocol state machine we need to implement. A leader can immediately reject
any responses that are not more current than its view (term). In turn, the
persistent log operates as a double ended queue, with entries either being
appended (or possibly deleted at followers) at the end and being deleted from
the front (after commit). This is critical to efficient operation of Cyclone, as
it allows a simple top level structure (a circular buffer of pointers) to
represent the log.

@@

The first key question we deal with is - how do Cyclone clients decide which
physical log to send a {\tt PUT} or {\tt GET} request to ? A key constraint here
is that all replicas should agree on the seqeunce of updates to a key - as
otherwise they would diverge in terms of state. Allowing updates to the same key
to go to different physical logs would lead to a divergence as there is no
guarantee about the order in which the independent RAFT instances would apply
the update. Therefore all updates to a key must go to the same physcial
log. A similar constraint applies to reads in order to ensure that any committed
writes are seen by readers. Allowing reads to a key to go to a different RAFT
instance than the write means that it could be served on a machine where the
write is yet to be applied but where the machine is the leader for the RAFT
instance. All reads and write to a key must therefore go to the same physical
log. We achieve this by hashing the key to select the physical log.

@@

Multiple physical logs in Cyclone operate in a shared nothing manner by
partitioning the NVM and SSD space evenly between them and by allocating
dedicated NIC queue pairs to each instance of RAFT. This works well because the
memory hierarchy including the NVM and the SSD efficiently support concurrent
operations. The only synchronization necessary is when doing reads or writes to
the single shared key value stores.  The level of concurrency therefore is
constrained only by the concurrency available in the software architecture of
the persistent key value store itself, for which a number of good designs
already exist~\cite{flodb}.

@@

\subsection{Ganged Operations}
Supporting reads or writes to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together in the face of failures. One option is to have an external
co-ordinator run a distributed transaction using two phase commit. In addition
to requiring extra external resources to host the co-ordinator one must also
worry about failure recovery. In addition this requires two round trips to the
leader to complete what should have been a single batched write request from a
client that is stateless and allowed to fail - such as with key value stores
like Rocksdb~\cite{rocksdb}. We therefore describe a simpler and more performant
solution called ganged operations integrated into Cyclone. This allows a client
to issue a batch of {\tt PUT} operations as a gang simultaneously and atomically
updating a batch of keys in a single call.

@@

Next, we need to simultaneously inject the ganged operation into all
participating physical logs on the machine hosting the leaders. Clients always
dispatch batched writes to a fixed co-ordinating physical log/RAFT instance
(henceforth called the co-ordinator), which is then responsible for forwarding
the request to the participating logs. We make use of packet cloning primitives
to avoid making a physical copy of the packet, generating indirect references
instead. The co-ordinator also adds a ``nonce'' - a unique timestamp to the
packet. In addition, the co-ordinator adds a unique view number to the packet
containing the term numbers of all the participating RAFT instances (read from
shared memory of the co-located leaders).  The event that applies a ganged
operation is described in the pseudocode of Figure~\ref{fig:ganged_ops}. We
defer the discussion of how we generate the nonce to later in this section. We
also assume that a unique barrier is allocated in shared memory for each ganged
operation. We discuss later how this is done without using dynamic allocation.

The complexity in Figure~\ref{fig:ganged_ops} arises from the need to handle
both shared memory concurrency and failure during replication. Without failure,
Figure~\ref{fig:ganged_ops} is straightforward. A barrier is executed on all
RAFT instances. Once replication is complete on all participating physical logs
(or this is a read ganged operation), a state indicated by all necessary bits
being set in the barrier mask, the RAFT instances simultaneously execute the
ganged operation.

Each RAFT instance always publishes the current term it is operating in. A
failure causes a new leader to be elected moving forward the term. If the
participants of a ganged operatation do not detect this case, they could be left
waiting for ever as ganged operation might never successfully replicate in the
failed over RAFT instance. We detect failure on any RAFT instance by having each
instance publish its current term and continuously comparing the view in the
ganged operation to it. If any RAFT instance has moved past that term, the
ganged operation is then terminated by having one of the still alive
participants set a failure flag and write the appropriate bit in the barrier
mask on behalf of the failed participant, moving execution past the barrier wait
- a technique we borrow from lock-free data structures~\cite{lock-free}.

\begin{figure*}
  \centering
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrrer.mask != mask of all participants
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

@@

Unless otherwise mentioned, we use a 60 byte header followed by an optional
payload for experiments. We log both the header and payload. Also, unless
otherwise mentioned we use 8 physical logs (and associated RAFT instances) each
mapped to a dedicated core. The remaining 32 cores are dedicated to the stub
server or Rocksdb, as the case may be.

@@

We now evaluate ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client - reflecting the
unloaded case - made ganged requests to the replicas. We varied the number of
physical log participating in the request from one to the full complement of 8.
Figure~\ref{fig:ganged} shows the results. Unloaded
latency increases slowly as we increase the number of active physical logs - to
around 32 us from the baseline of 21 us. The reason for this increase is that
the DPDK userspace NIC driver pipelines request processing for communication
between the CPU core and the NIC. Simultaneous replicaton on multiple physical
logs represents a worst case for this arrangement as all eight cores try to
send replication messages to their replicas, adding about a microsecond of
serial latency each, to dispatch requests to the NIC. Regardless, the experiment
underlines the value of our ganged operation design over a distributed
transaction that would need 4 network hops for the two phase commit (and
replication) using an external co-ordinator, a minimum of 40us.

@@

We now evaluate Cyclone integrated with the Rocksdb persistent key value
store. Rocksdb is a complex industrial strength persistent key value store and
an important target for practical use of Cyclone.  Rocksdb is accompanied by a
complex array of performance tuning knobs to get the best performance from
flash. Designing LSM tree based key value stores to extract good performance
from flash and make best use of DRAM on the machine is an area of active
research~\cite{flodb, triad}. We wished to make sure that our conclusions would
apply even as Rocksdb performance improves with the integration of new
ideas. Keeping this in mind, we configured RocksDB to place all files for the
key value store (SSTables) on a RAMdisk, which presumably represents the limit
in performance for both software enhancements as well as improvements to
flash-based SSDs. However - both Rocksdb's own write ahead log and the
alternative of Cyclone's second level flashlog - are placed on the SSDs. In
effect we therefore make availability a harder problem in this setting,
guaranteeing our design can handle the immediate future.

We use 8 bytes keys with either 8 bytes values or 256 byte values.
Since our focus is replication and only update requests are replicated, we use a
pure update workload to test the capability of the system.

@@

The performance of Rocksdb with Cyclone for the small update workload is shown
in Figure~\ref{fig:rocksdb} - essentially presenting the solution to the problem
description graph in Figure~\ref{fig:problem}.  We consider four different
setups. The line labeled 'rocksdb' is the key value store running with no
logging whatsover - a system crash would lead to data loss. The line labeled
'rocksdb/wal' is for Rocksdb running with its write ahead logging turned on. The
large gap between these two is the overhead of the existing Rocksdb WAL
solution. The line labeled 'rocksdb/Cyclone 1 way' is a two level Cyclone log
but without any replication. The line almost exactly tracks the performance of
Rockdb. As suggested by the baseline replication performance, Cyclone is able to
provide a write ahead log with no overhead to Rocksdb. The line labeled
'rocksdb/Cyclone 3 way' is with 3-way replication turned on. Other than a 20us
delta due to the extra network round trip, the line almost exactly tracks
Rocksdb performance with no logging. Cyclone therefore provides high
availability to Rocksdb at a fraction of the cost of its existing single machine
write ahead log. We also repeat the experiment for the larger update size in
Figure~\ref{fig:rocksdb_256}. The conclusions are identical: Cyclone solves
Rocksdb's write ahead logging problem.

@@

DARE~\cite{dare} implements state machine replication over RDMA, using a custom
replication protocol. DARE does not consider persistence in its design. It is
worthwhile to point out that since DARE allows any replica to become the leader
it requires a log adjustment protocol for the leader to bring its logs up to
date, a situation we avoid by using RAFT as described in
Section~\ref{sec:dm}. Using 40 Gb/s speeds, and 3-way replication DARE reported
being able to replicate 500K requests/sec. This is without persisting the log of
updates. In contrast, Cyclone replicates upwards of 6 million similarly sized
requests each second with 4*10 Gbps ethernet (without the second level
flashlog). DARE reports an unloaded write latency of 15us, which is lower than
the unloaded write latency of Cyclone at 20us, thanks to infiniband. The
conclusions we can draw are similar to the comparison with Mojim: a carefully
implemented log replication system obviates the need for network offload when
doing replication.

>>>>

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{387}

@@

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. This is due to the fixed cost of a round trip to the SSD through the IO
interface (NVMe in our case), regardless of the volume of data being
synchronously persisted. Small updates (sum of key and value data) are
surprisingly common in observed workloads for key value pairs. For example,
Facebook~\cite{fb_workload_analysis} report that 90\% of the space allocated in
key value stores is for data items under 500 bytes in size.

Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are significantly more expensive than standard DIMMs. This means that they will
likely be available in small quantities. We compensate for this by periodically
draining the NVDIMM log into a log placed on a standard SSD on the system.

The second key contribution of this work is in showing that the throughput
bottlenecks in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum - a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory eliminating block IO from the persistence step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

The rest of the paper is organized as follows. We describe Cyclone's system
architecture in Section~\ref{sec:sysarch}. We describe the two level log
structure in Section~\ref{sec:storage}. The set of optimizations to the network
stack is decribed in Section~\ref{sec:network}. Both the storage (NVM, SSD) as
well as the NIC expose significant parallelism that is left unused when Cyclone
is used to replicate a single log. To exploit this parallelism we use Cyclone to
replicate multiple logs, each using a different instance of the RAFT consensus
protocol. We show how to exploit the commutativity inherent in key-value store
interfaces to make use of this parallelism in Cyclone's design in
Section~\ref{sec:parallelism}.  A detailed evaluation of Cyclone's replication
performance is provided in the context of RocksDB - a popular persistent key
value store - in Section~\ref{sec:evaluation}. We then discuss related work
before concluding.

@@

\begin{itemize}
\item GET(K): returns value corresponding to key K
\item PUT(K, V): sets the value of key K to value V
\item DELETE(K): deletes the key
\end{itemize}

A key value store typically serves multiple client requests concurrently and is
linearizable~\cite{linearizability}. This means any history of requests and
responses can be rearranged such that responses immediately follow corresponding
requests and if a response precedes a request in the original history it also
does so in the rearranged history. Key value stores such as RocksDB also
support atomic writes ({\tt PUT} operations) to multiple keys, called batched
writes and the ability to take snapshots.

@@

Cyclone uses RAFT~\cite{raft} to keep the logs in sync across the replicas. On
failure of the leader replica, a new leader is automatically elected - RAFT
ensures that the new leader has the most up to date log. The Cyclone client
library automatically locates the new leader on a failover - clients therefore
see only a brief interruption in service. Although read requests (GET
operations) are not logged, they are serviced in FIFO order with the write
requests (PUT/DELETE operations) currently in Cyclone's log at the moment the
read request is received. This ensures linearizability in the face of failovers.

@@

The two level log design in Cyclone lays the foundation for improvements to the
network stack that we discuss next. We use RAFT to replicate the NVM log from
the leader replica to the followers. Each replica \emph{independently} drains
committed (in terms of RAFT consensus) entries from the NVM log to the
flashlog. Block IO therefore is no longer a consideration when optimizing data
movement over the network, unlike systems that need to deal with such
problems~\cite{reflex}. Replication reduces to efficiently sending a block of
data already present in directly attached memory over the network.

@@

The core network operation in replication is to receive a request from the
client at the leader replica and send that exact same request (as a log entry)
to all follower replicas. The leader therefore \emph{multicasts} a received
request packet to follower replicas, a problem well studied in the networking
community when building software packet switches. Software packet switches reach
impressive speeds of millions of packets forwarded per second~\cite{dpdk_perf}, a
number far in excess of the few hundreds of packets we manage in
Figure~\ref{fig:problem}. The key insight in removing network overheads in
Cyclone is therefore to approximate a software packet switch for the networking
component of Cyclone.

This is however not a straightforward mapping. Software packet switches make a
simple (usually stateless or sometimes flow associated) decision for each
packet. In contrast, with Cyclone the role of replicas is asymmetric. Log
entries flow from leader replica to followers and acknowledgments flow
back. Further, the assigned role to a replica can change on reconfiguration
after a failure. A replica must also reject packets from old configurations. The
action associated with a received packet therefore is far more complex than
simply appending it to a log.  The \emph{challenge} is to do this efficiently in
software.

@@

\subsection{Memory Management}
\label{sec:mm}
We direct DPDK to use pages backed by NVM for packet buffers. DPDK uses a
concurrent memory allocator based on reference counting - used by both the NIC
driver and CPU cores. This means that packet allocation and deallocation does
not happen in the same order as their corresponding position in the replicated
RAFT log. To deal with this, we use a level of indirection as shown in
Figure~\ref{fig:nvm_log}. The NVM log is maintained as a circular log of fixed
sized pointers to the actual packets. Adding a level of indirection in the NVM
log allows us to separate the FIFO ordered circular log being manipulated by
RAFT from packet data being managed by the memory allocator of DPDK. Both the
circular log and packet data are in NVM. An advantage of this scheme is that it
makes recovery from NVM easy - appends to the circular pointer log are atomic
and we can use the pointer log to recover allocator state i.e. what pieces of
NVM are currently in use by the log.

@@

A key advantage of RAFT is that unlike alternatives such as
MultiPaxos~\cite{multipaxos} or Quorum based replication~\cite{quorum}, the
leader is guaranteed to have the most up to date logs. There is therefore no
case where a leader needs to receive log entries to ``fill'' holes in its logs
from follower replicas, simplifying the protocol state machine we need to
implement. A leader can immediately reject any responses that are not more
current than its view (term). In turn, the persistent log operates as a double
ended queue, with entries either being appended (or possibly deleted at
followers) at the end and being deleted from the front (after commit). This is
critical to efficient operation of Cyclone, as it allows a simple top level
structure (a circular buffer of pointers) to represent the log.

\subsection{Data Movement}
\label{sec:dm}
Software packet switches built on DPDK try to touch as little of the packet data
as possible, minimizing movement of data up and down the cache and memory
hierarchy. Log replication looks like multicast that does not require deep
packet inspection. Software packet switches implementing multicast therefore
simply manipulate packet headers to produce new packets to send on the output
ports. We designed our implementation of Raft around the same principle.

@@

We now turn our attention to the persistence step in
Figure~\ref{fig:control_plane}. RAFT requires that the log entry be persisted
\emph{before} it is multicast out to follower replicas. DPDK userspace NIC
drivers operate in DDIO mode~\cite{ddio} where the packet is directly written
into the CPU cache rather than first being DMAed into DRAM and then fetched by
the CPU on demand. Since the NVM is directly attached we need to persist the
packet to it by executing a cacheline flush ({\tt clflush}) instruction for
every cacheline in the packet and the pointer in the pointer buffer to persist
these via the memory bus. This is not too onerous a burden because we can use
the newly introduced {\tt clflush-opt}~\cite{clflush_opt} instruction
specifically intended to efficiently flush to persistent memory without the
overhead of the serialization normally introduced by {\tt clflush}. This allows
us to hit full memory bandwidth on present generation platforms, a quantity in
excess of 200 Gb/s per core, well above the near term speeds of network
interface cards. We execute a single serializing {\tt sfence} before the
sequence of {\tt clflushes} to make sure any dirty cacheline due to header
manipulation related writes from the CPU are sent to cache.

\subsection{Batching}

@@

@@

The first question we deal with is - how do Cyclone clients decide which
physical log to send a request to ? A key constraint here is that all replicas
should agree on the seqeunce of updates to a key - as otherwise they would
diverge in terms of state. Allowing updates to the same key to go to different
physical logs would lead to a divergence as there is no guarantee about the
order in which the independent RAFT instances would apply the update. Therefore
all updates to a key must go to the same physcial log. A similar constraint
applies to reads in order to ensure that all committed writes are seen by
readers. Allowing reads to a key to go to a different RAFT instance than the
write means that it could be served on a machine where the write is yet to be
applied. All reads and write to a key must therefore go to the same physical
log. We achieve this by hashing the key to select the physical log.

@@

Multiple physical logs in Cyclone operate in a shared nothing manner by
partitioning the NVM and SSD space evenly between them and by allocating
dedicated NIC queue pairs to each instance of RAFT. This works well because the
memory hierarchy including the NVM and the SSD efficiently support concurrent
operations. The only synchronization necessary is when doing reads or writes to
the single shared key value store.  The level of concurrency therefore is
constrained only by the concurrency available in the software architecture of
the persistent key value store itself, for which good designs
exist~\cite{flodb}.

@@

\subsection{Ganged Operations}
Supporting reads or writes to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together even in the face of failures. One option is to have an
external co-ordinator run a distributed transaction using two phase commit. In
addition to requiring extra external resources to host the co-ordinator one must
also worry about failure recovery. This also requires two round trips to the
leader to complete what should have been a single batched write request from a
client that is stateless and allowed to fail - such as with key value stores
like Rocksdb~\cite{rocksdb}. We therefore describe a simpler and more performant
solution called ganged operations integrated into Cyclone. This allows a
stateless client to issue a batch of update operations as a gang, simultaneously
and atomically updating a batch of keys in a single call.

@@

Next, we need to simultaneously inject the ganged operation into all
participating physical logs on the machine hosting the leaders. Clients always
dispatch batched writes to a fixed co-ordinating physical log/RAFT instance
(henceforth called the co-ordinator), which is then responsible for forwarding
the request to the participating logs. We make use of packet cloning primitives
to avoid making physical copies of the packet, generating indirect references
instead. The co-ordinator also adds a ``nonce'' - a unique timestamp to the
packet. In addition, the co-ordinator adds a unique view number to the packet
containing the term numbers of all the participating RAFT instances (read from
shared memory of the co-located leaders).  The event that applies a ganged
operation is described in the pseudocode of Figure~\ref{fig:ganged_ops}. We
defer the discussion of how we generate the nonce to later in this section. We
also assume that a unique barrier is allocated in shared memory for each ganged
operation. We discuss later how this is done without using dynamic allocation.

The complexity in Figure~\ref{fig:ganged_ops} arises from the need to handle
both shared memory concurrency and failure during replication. Without failure,
Figure~\ref{fig:ganged_ops} is straightforward. A barrier is executed on all
RAFT instances. Once replication is complete on all participating physical logs,
a state indicated by all necessary bits being set in the barrier mask, the RAFT
instances simultaneously execute the ganged operation.

A failure causes a new leader to be elected for the affected physical log,
moving forward the term. If the participants of a ganged operatation do not
detect this case, they could be left waiting forever. We detect failure on any
RAFT instance by having each instance publish its current term and continuously
comparing the view in the ganged operation to it. If any RAFT instance has moved
past that term, the ganged operation is then terminated.

\begin{figure}
  \centering
  \scriptsize
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrier.mask != mask of all participants
        if barrier.failed
         send retry to client
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         send retry to client
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure}

@@

Unless otherwise mentioned, we use a 60 byte header followed by an optional
payload for experiments. We log both the header and payload. Also, unless
otherwise mentioned we use 8 physical logs (and associated RAFT instances) each
mapped to a dedicated core. The remaining cores are dedicated to the stub
server or Rocksdb, as the case may be.

@@

We now evaluate ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client - reflecting the
unloaded case - made ganged requests to the replicas. We varied the number of
physical log participating in the request from one to the full complement of 8.
Figure~\ref{fig:ganged} shows the results. Unloaded
latency increases slowly as we increase the number of active physical logs - to
around 32 us from the baseline of 21 us. The reason for this increase is that
the DPDK userspace NIC driver pipelines request processing for communication
between the CPU core and the NIC. Simultaneous replicaton on multiple physical
logs represents a worst case for this arrangement as all eight cores try to
send replication messages to their replicas, adding about a microsecond of
serial latency each, to dispatch requests to the NIC. Regardless, the experiment
underlines the value of our ganged operation design over a distributed
transaction that would need 8 network hops for the two phase commit (and
replication) using an external co-ordinator, a minimum of 80us.

@@

Next, we evaluate Cyclone integrated with the Rocksdb persistent key value
store. Rocksdb is a complex industrial strength persistent key value store and
an important target for practical use of Cyclone.  Rocksdb is accompanied by a
complex array of performance tuning knobs to get the best performance from
flash. Designing LSM tree based key value stores to extract good performance
from flash and make best use of DRAM on the machine is an area of active
research~\cite{flodb, triad}. We wished to make sure that our conclusions would
apply even as Rocksdb performance improves with the integration of new
ideas. Keeping this in mind, we configured RocksDB to place all files for the
key value store (SSTables) on a RAMdisk, which presumably represents the limit
in performance for both software enhancements as well as improvements to
flash-based SSDs. However - both Rocksdb's own write ahead log and the
alternative of Cyclone's second level flashlog - are placed on the SSDs. In
effect we therefore make availability a harder problem in this setting,
guaranteeing our design can handle the immediate future.

We use 8 byte keys with either 8 byte values or 256 byte values.  Since our
focus is replication and only update requests are replicated, we use a pure
update workload to test the capability of the system. This involves loading 100
million key value pairs and during the test updating the value associated with a
key picked at random.

@@

The performance of Rocksdb with Cyclone for the small update workload is shown
in Figure~\ref{fig:rocksdb} - essentially presenting the solution to the problem
description graph in Figure~\ref{fig:problem}.  We consider four different
setups. The line labeled 'rocksdb' is the key value store running with no
logging whatsover - a system crash would lead to data loss. The line labeled
'rocksdb/wal' is for Rocksdb running with its write ahead logging turned on. The
large gap between these two is the overhead of the existing Rocksdb WAL
solution. The line labeled 'rocksdb/Cyclone 1 way' is a two level Cyclone log
but without any replication. The line almost exactly tracks the performance of
Rockdb. Cyclone is able to provide a write ahead log with no overhead to
Rocksdb. The line labeled 'rocksdb/Cyclone 3 way' is with 3-way replication
turned on. Other than a 20us delta due to the extra network round trip, the line
almost exactly tracks Rocksdb performance with no logging. Cyclone therefore
provides high availability to Rocksdb at a fraction of the cost of its existing
single machine write ahead log. We also repeat the experiment for the larger
update size in Figure~\ref{fig:rocksdb_256}. The conclusions are identical:
Cyclone solves Rocksdb's write ahead logging problem.

@@

DARE~\cite{dare} implements state machine replication over RDMA, using a custom
replication protocol. DARE does not consider persistence in its design. It is
worthwhile to point out that since DARE allows any replica to become the leader
it requires a log adjustment protocol for the leader to bring its logs up to
date, a situation we avoid by using RAFT as described in
Section~\ref{sec:mm}. Using 40 Gb/s speeds, and 3-way replication DARE reported
being able to replicate 500K requests/sec. This is without persisting the log of
updates. In contrast, Cyclone replicates upwards of 6 million similarly sized
requests each second with 4*10 Gbps ethernet (without the second level
flashlog). DARE reports an unloaded write latency of 15us, which is lower than
the unloaded write latency of Cyclone at 20us, thanks to infiniband. The
conclusions we can draw are similar to the comparison with Mojim: a carefully
implemented log replication system obviates the need for network offload when
doing replication.
